State and Parameter Estimation for Spatio-Temporal Neural Fields

### 2010-05-05 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex

% 
% Due to the high dimensionality of the brain and the current electrode systems, we can not expect to have more sensors than basis functions.
\section{State and Parameter Estimation}\label{StateAndParameterEstimationSection} In this section we describe  estimating the states that scale the field basis functions, $\mathbf{x}_t$, the connectivity kernel parameters, $\boldsymbol \theta$, and the synaptic time constant, $\xi$. The estimation process consists of a two part iterative algorithm, consisting of the state estimation step followed by the parameter estimation step. At each iteration, a sequence of estimated states is used to calculate the parameter set and the resulting parameters is used to generate a new sequence of states for the next iteration. The procedure stops when the algorithm converges. The algorithm is initialized using a bounded random state sequence which guarantees the initial parameter set forms a stable kernel.

The state estimation technique uses the Unscented Rauch Tung Striebel Smoother (URTSS)~\cite{Sarkka2010}. The URTSS uses the Unscented Kalman Filter (UKF)~\cite{Julier1997,Merwe2003} in the forward iteration to estimate posterior states, $\hat{\mathbf x}_t^{f}$, followed by a backward iteration to compute the smoothed state estimates, $\hat{\mathbf x}_t^{b}$. The first and the second order moments of the predicted state is captured by propagating the so-called sigma points through the state equation. The sigma points $\mathcal X_i$ are calculated using the unscented transform as follows

>>>>

% 
% Due to the high dimensionality of the brain and the current electrode systems, we can not expect to have more sensors than basis functions.
\section{State and Parameter Estimation}\label{StateAndParameterEstimationSection} In this section we describe  estimating the states that scale the field basis functions, $\mathbf{x}_t$, the connectivity kernel parameters, $\boldsymbol \theta$, and the synaptic time constant, $\xi$. The estimation process consists of a two part iterative algorithm, consisting of the state estimation step followed by the parameter estimation step. At each iteration, a sequence of estimated states is used to calculate the parameter set and the resulting parameters is used to generate a new sequence of states for the next iteration. The procedure stops when the algorithm converges. The algorithm is initialized using a bounded random state sequence which guarantees the initial parameter set forms a stable kernel.
\label{StateAndParameterEstimationSection}

In this section we describe the procedure for estimating the states that scale the field basis functions, $\mathbf{x}_t$, the connectivity kernel parameters, $\boldsymbol \theta$, and the synaptic time constant, $\xi$. The estimation process consists of a two part iterative algorithm, consisting of the state estimation step followed by the parameter estimation step. At each iteration, the state estimate sequence is used to calculate the parameter set and the resulting parameters are used to generate a new state sequence for the next iteration. The procedure stops when the algorithm converges. The algorithm is initialized using a bounded random state sequence which guarantees the initial parameter set forms a stable kernel.

The state estimation technique uses the Unscented Rauch Tung Striebel Smoother (URTSS)~\cite{Sarkka2010}. The URTSS uses the Unscented Kalman Filter (UKF)~\cite{Julier1997, Merwe2003} in the forward iteration to estimate posterior states, $\hat{\mathbf x}_t^{f}$, followed by a backward iteration to compute the smoothed state estimates, $\hat{\mathbf x}_t^{b}$. The first and the second order moments of the predicted state is captured by propagating the so-called sigma points through the state equation. The sigma points $\mathcal X_i$ are calculated using the unscented transform as follows
